---
title       : Fiji Earthquake Analysis
subtitle    : Coursera Data Science Project
author      : Yan Wang
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Instruction

This presentation is part of the Course Project for the Coursera Developing Data Products class. For the first part, a Shiny application is created and deployed on Rstudio's server. For the second part, a reproducible pitch presentation about the application is created using slidify.

The app developed for the first part of the assignment is avalilable at:

https://yanw.shinyapps.io/FiJiEarthquakeAnalysis/

Source code for ui.R and server.R files are available on the GitHub:

https://github.com/yan10/datasciencecoursera/tree/master/DataProduct/shiny

--- .class #id 

## Shiny Application

This Shiny application helps you analyzing the location, depth and magnitude of earthquakes off Fiji, using the quakes dataset from [R]. It is able to filter out earthquakes depanding on criterias you set And visualize different earthquakes on map. 

The data set give the locations of 1000 seismic events of MB > 4.0. The events occurred in a cube near Fiji since 1964.

```{r, echo = FALSE}
data("quakes")
head(quakes)
```

--- .class #id 

## Locations of Earthquakes off Fiji

First, you need to set the minimum threshold of number of stations reporting the earthquakes. These information will be used to filter out earthquakes don't meet the requirement.

Second, you can drag the slider bar to select the range of earthquake depth and earthquake maginitude you want to include.

After you set the criterias, this application will show you a map of Fiji with circles stand for each earthquake, the radius of circle stands for different value. You could choose to see two differenent types of map: OpenStreetMap or MapQuestOpen.Aerial. Also, you could choose to see different types of reported data: Depth or Magnitude or Number of Stations Reporting. Also you could put cursor on the circle and it will give you corresponding information.

--- .class #id 

## Earthquake Map

Here is the code to generate a basic map about Fiji earthquake. The actual applicaton has more features.

```{r, warning=FALSE,comment=FALSE,results='hide',message=FALSE}
require(leaflet)
data("quakes")
map <- leaflet(height = 650, width = 900)
map <- addTiles(map, group = "OpenStreetMap")
map <- addCircles(map,
                  lng = quakes$long,
                  lat = quakes$lat,
                  radius = 5 ^ quakes$mag,
                  popup = paste0('Magnitude: ', quakes$mag),
                  group = "Magnitude",
                  fillColor = "green",
                  fill = TRUE,
                  stroke = FALSE,
                  fillOpacity = 0.5
                  )
```

